\chapter{Performance Evaluation}
\label{cha:evaluation}
\vspace{0.4 cm}

In this chapter, the proposed system is validated and the performance of the models for the different use cases are evaluated.
The first section presents the datasets provided by MIWEnergia.
Subsequently, the adapted evaluation methodology is described.
Finally, the evaluation of the performance of the models for the different use cases is presented.
After this chapter, it will be clear how the system has been validated and what the performance achieved by the proposed system are.


\section{MIWEnergia datasets}
\label{sec:datasets}
\vspace{0.2 cm}

Describe the MIWEnergia datasets ...

- Aggregated consumption data: from ... to ..., horly aggregated, ... total entries.

- Consumption data of 4 customers: from ... to ..., horly aggregated, ... total entries.

- Production data of 8 palnts: from ... to ..., horly aggregated, ... total entries.

Report plots of the data ...


\section{Evaluation methodology}
\label{sec:methodology}
\vspace{0.2 cm}

General description of the evaluation methodology ...

Describe the metrics (MAPE e MAE) ...

Describe the test of the performance of the models ...

Treat block validation: describe blocked k-fold e hold-out on more periods ...


\section{Electricity demand forecasting}
\label{sec:demandval}
\vspace{0.2 cm}

Analyze the data (descriptive analytics) and the results on the electricity demand forecasting task (predictive analytics) ...

Data stats and correlation with weather data ...

Describe the choice of parameters for models ... (list with motivations)

Summary table with results ...


\section{Consumption baseline forecasting}
\label{sec:baselineval}
\vspace{0.2 cm}

Analyze the data (descriptive analytics) and the results on the consumption baseline forecasting task (predictive analytics) ...

Data stats and correlation with weather data ...

Describe the choice of parameters for models ... (list with motivations)

Summary table with results ...


\section{Electricity production forecasting}
\label{sec:productionval}
\vspace{0.2 cm}

Analyze the data (descriptive analytics) and results on the electricity production forecasting task (predictive analytics) ...

Data stats and correlation with weather data ...

Describe the choice of parameters for models ... (list with motivations)

Summary table with results ...
